---
title: "Two Variables: Both Quantitative"
author: Orville D. Hombrebueno
date: today
format:
  html:
    embed-resources: true
editor: visual
toc: true
code-link: true
code-fold: true
cache: true
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# install packages: install.packages("package")
# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)
library(broom)

# load data
data(mtcars)
```

## Motor Trend Car Road Tests

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

A data frame with 32 observations on 11 (numeric) variables.

-   `mpg` Miles/(US) gallon

-   `cyl` Number of cylinders

-   `disp` Displacement (cu.in.)

-   `hp` Gross horsepower

-   `drat` Rear axle ratio

-   `wt` Weight (1000 lbs)

-   `qsec` 1/4 mile time

-   `vs` Engine (0 = V-shaped, 1 = straight)

-   `am` Transmission (0 = automatic, 1 = manual)

-   `gear` Number of forward gears

-   `carb` Number of carburetors

The first 6 rows of the data set are shown in @tbl-mtcars

```{r}
#| label: tbl-mtcars
#| warning: false
#| message: false
#| error: false
#| tbl-cap: "The first 6 observations of the data set."

mtcars |> # call the data
  head() |> # get the first six rows of the data
  kable() # create a table
```

## Descriptive Statistics

The mean, standard deviation, minimum, and maximum of `hp` and `mpg` in the `mtcars` data set are shown in @tbl-summary. There is a total of 32 observations.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of the data set."
#| message: false

mtcars |> # call the data
  tbl_summary( # create a table
    include = c(mpg, hp), # select variables
    statistic = all_continuous() ~ "{mean} ± {sd} ({min} - {max})", # descriptive statistics
    digits = all_continuous() ~ 2 # no. of digits (rounding)
  )
```

# Hypothesis Testing

## Research Question

Is there a significant relationship between gross horsepower and miles/(US) gallon?

## Hypothesis and Test Statistics

-   $H_0:\rho=0$; there is no linear relationship between gross horsepower and miles/(US) gallon.

-   $H_a:\rho\ne0$; there is a linear relationship between gross horsepower and miles/(US) gallon.

-   Significance level, $\alpha=5\%$.

-   **Pearson correlation**

-   $n>30$; large sample.

## Results and Interpretation

The scatter plot and test statistics is shown on @fig-test.

```{r}
#| label: fig-test
#| fig-cap: Scatter plot and test statistics.
#| fig-height: 3
#| warning: false
#| error: false
#| message: false

ggscatterstats(
  mtcars, hp, mpg,
  xlab = "Gross Horsepower",
  ylab = "Miles/(US) Gallon",
  type = "parametric"
) +
  labs(caption = NULL) # remove caption
```

Based on the scatter plot and the *p*-value being less than 5% (*p*-value = .001), we reject the null hypothesis. There is strong evidence of a linear relationship between gross horsepower and miles/(US) gallon. The correlation ($\rho=-0.78$) suggests a moderate negative relationship between the two variables. Increased gross horsepower is moderately associated with decreased miles/(US) gallon.
