---
title: "Two Variables: Quantitative and Qualitative"
author: Orville D. Hombrebueno
date: today
format:
  html:
    embed-resources: true
editor: visual
toc: true
code-link: true
code-fold: true
cache: true
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# install packages: install.packages("package")
# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)
library(broom)
library(car)

# load data
data(mtcars)
```

## Motor Trend Car Road Tests

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

A data frame with 32 observations on 11 (numeric) variables.

-   `mpg` Miles/(US) gallon

-   `cyl` Number of cylinders

-   `disp` Displacement (cu.in.)

-   `hp` Gross horsepower

-   `drat` Rear axle ratio

-   `wt` Weight (1000 lbs)

-   `qsec` 1/4 mile time

-   `vs` Engine (0 = V-shaped, 1 = straight)

-   `am` Transmission (0 = automatic, 1 = manual)

-   `gear` Number of forward gears

-   `carb` Number of carburetors

The first 6 rows of the data set are shown in @tbl-mtcars

```{r}
#| label: tbl-mtcars
#| warning: false
#| message: false
#| error: false
#| tbl-cap: "The first 6 observations of the data set."

mtcars |> # call the data
  head() |> # get the first six rows of the data
  kable() # create a table
```

## Descriptive Statistics

The mean, standard deviation, minimum, and maximum of `mpg` when grouped according to `am` are shown in @tbl-summary. There is a total of 32 observations.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics."
#| message: false

mtcars |> # call the data
  mutate(
    am = factor(
      am, levels = c(0, 1),
      labels = c("Automatic", "Manual")
    )
  ) |>
  tbl_summary( # create a table
    by = am, # grouping
    include = c(mpg, am), # select variables
    statistic = all_continuous() ~ "{mean} ± {sd} ({min} - {max})", # descriptive statistics
    digits = all_continuous() ~ 2 # no. of digits (rounding)
  )
```

# Hypothesis Testing

### Research Question

Is there a significant difference in miles/(US) gallon when grouped according to transmission?

### Hypothesis and Test Statistics

-   $H_0:\mu_A=\mu_M$; the mean miles/(US) gallon is the same when grouped according to transmission.

-   $H_a:\mu_A\ne\mu_M$; the mean miles/(US) gallon is different when grouped according to transmission.

-   Significance level, $\alpha=5\%$.

-   **Welch's t-test for 2 independent samples**

-   Unknown and unequal variance (see @tbl-levene).

```{r}
#| label: tbl-levene
#| tbl-cap: Levene's test for homogeneity of variance.

mtcars_levene <- mtcars |> # call the data
  mutate(
    am = factor(
      am, levels = c(0, 1),
      labels = c("Automatic", "Manual")
    )
  )

leveneTest(mpg ~ am, data = mtcars_levene, center =  mean) |>
  tidy() |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  kable()
```

-   Both samples follow a normal distribution (see @tbl-shapiro).

```{r}
#| label: tbl-shapiro
#| tbl-cap: Shapiro-Wilk normality test.

mtcars_shapiro_m <- mtcars_levene |>
  filter(am == "Manual") |>
  select(mpg)

shapiro_m <- shapiro.test(mtcars_shapiro_m$mpg) |>
  tidy() |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))

mtcars_shapiro_a <- mtcars_levene |>
  filter(am == "Automatic") |>
  select(mpg)

shapiro_a <- shapiro.test(mtcars_shapiro_a$mpg) |>
  tidy() |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))

bind_rows(shapiro_a, shapiro_m) |>
  mutate(
    transmission = c("Automatic", "Manual")
  ) |>
  relocate(transmission, .before =1) |>
  kable()

```

### Results and Interpretation

The plot and test statistics is shown on @fig-stats.

```{r}
#| label: fig-stats
#| fig-cap: Comparison of `mpg` by `am`.
#| fig-height: 3

mtcars |> # call the data
  mutate(
    am = factor(
      am, levels = c(0, 1),
      labels = c("Automatic", "Manual")
    )
  ) |>
  ggbetweenstats(
    x = am,
    y = mpg,
    plot.type = "box", # for boxplot
    type = "parametric", # for student's t-test
    var.equal = FALSE, # equal variances
    centrality.plotting = TRUE # remove mean
  ) +
  labs(caption = NULL) # remove caption
```

Based on the plot and the *p*-value being less than 5% (*p*-value = .001), we reject the null hypothesis. There is strong evidence that the mean miles/(US) gallon is different when grouped according to transmission. Cars with manual transmission tend to have a higher mean miles/(US) gallon. Thus, manual cars are more fuel efficient compared to automatic cars.
